Pangenome graph layout by Path-Guided Stochastic Gradient Descent

### 2023-10-13 manuscript/main.tex
add feedback from Sven

	\author[Heumos, Guarracino \textit{et~al}.]{
        Simon~Heumos\,$^{\orcidlink{0000-0003-3326-817X}1,2,\dagger}$,
        Andrea~Guarracino\,$^{\orcidlink{0000-0001-9744-131X}3,4,\dagger}$,
        Jan-Niklas M. Schmelzle\,$^{\orcidlink{0000-0001-8566-4049}5,6}$,
        Jiajie Li\,$^{\orcidlink{0000-0002-6775-2843}6}$,
        Zhiru Zhang\,$^{\orcidlink{0000-0002-0778-0308}6}$,
        Jörg Hagmann\,${\orcidlink{0000-0002-7232-3103}^7}$,
        Sven Nahnsen\,$^{\orcidlink{0000-0002-4375-0691}1,2}$,
        Pjotr Prins\,$^{\orcidlink{0000-0002-8021-9162}3}$,
        Erik~Garrison\,$^{\orcidlink{0000-0003-3821-631X}\text{\sfb 3},*}$
    }

    \address{
        $^1$Quantitative Biology Center (QBiC), University of Tübingen, Tübingen 72076, Germany \\
        $^2$Biomedical Data Science, Department of Computer Science, University of Tübingen, Tübingen 72076, Germany \\
        $^3$Department of Genetics, Genomics and Informatics, University of Tennessee Health Science Center, Memphis, TN 38163, USA \\
        $^4$Genomics Research Centre, Human Technopole, Milan 20157, Italy \\
        $^5$Department of Computer Engineering, School of Computation, Information and Technology (CIT), Technical University of Munich, Munich 80333, Germany \\
        $^6$School of Electrical and Computer Engineering, Cornell University, Ithaca, NY 14853, USA \\
        $^7$Computomics GmbH, Eisenbahnstr. 1, 72072 Tübingen, Germany \\
    }

@@

	\section{Introduction}
	Reference genomes are widely used in genetics, serving as a foundation for a variety of analyses, including gene annotation, read mapping, and variant detection \citep{Singh2022}.
	However, this linear model is becoming obsolete given the accessibility to hundreds or even thousands of high-quality genomes.
	A single genome can not fully represent the genetic diversity of any species, resulting in reference bias \citep{Ballouz2019}.
	In contrast, a pangenome models the entire set of genomic elements of a given population \citep{Tettelin_2008,cpang2018,Eizenga_2020, Sherman_2020}.
	Pangenomes can be represented as a sequence graph incorporating sequences as nodes and their relationships as edges \citep{Hein1989}.
	%These nodes are shared for identical sequences, such as homologs, paralogs, and orthologs.
	In the variation graph model \citep{Garrison:2018}, genomes are encoded as paths traversing the nodes in the graph.
	%A bidirected graph can even contain both strands of DNA as paths. \\

>>>>

	\author[Heumos, Guarracino \textit{et~al}.]{
        Simon~Heumos\,$^{\orcidlink{0000-0003-3326-817X}1,2,\dagger}$,
        Andrea~Guarracino\,$^{\orcidlink{0000-0001-9744-131X}3,4,\dagger}$,
        Jan-Niklas M. Schmelzle\,$^{\orcidlink{0000-0001-8566-4049}5,6}$,
        Jiajie Li\,$^{\orcidlink{0000-0002-6775-2843}6}$,
        Zhiru Zhang\,$^{\orcidlink{0000-0002-0778-0308}6}$,
        Jörg Hagmann\,${\orcidlink{0000-0002-7232-3103}^7}$,
        Sven Nahnsen\,$^{\orcidlink{0000-0002-4375-0691}1,2,8}$,
        Pjotr Prins\,$^{\orcidlink{0000-0002-8021-9162}3}$,
        Erik~Garrison\,$^{\orcidlink{0000-0003-3821-631X}\text{\sfb 3},*}$
    }

    \address{
        $^1$Quantitative Biology Center (QBiC), University of Tübingen, Tübingen 72076, Germany \\
        $^2$Biomedical Data Science, Department of Computer Science, University of Tübingen, Tübingen 72076, Germany \\
        $^3$Department of Genetics, Genomics and Informatics, University of Tennessee Health Science Center, Memphis, TN 38163, USA \\
        $^4$Genomics Research Centre, Human Technopole, Milan 20157, Italy \\
        $^5$Department of Computer Engineering, School of Computation, Information and Technology (CIT), Technical University of Munich, Munich 80333, Germany \\
        $^6$School of Electrical and Computer Engineering, Cornell University, Ithaca, NY 14853, USA \\
        $^7$Computomics GmbH, Eisenbahnstr. 1, 72072 Tübingen, Germany \\
        $^8$M3 Research Center, University Hospital Tübingen, 72076 Tübingen, Germany \\
    }

@@

	\section{Introduction}
	Reference genomes are widely used in genomics, serving as a foundation for a variety of analyses, including gene annotation, read mapping, and variant detection \citep{Singh2022}.
	However, this linear model is becoming obsolete given the accessibility to hundreds or even thousands of high-quality genomes.
	A single genome can not fully represent the genetic diversity of any species, resulting in reference bias \citep{Ballouz2019}.
	In contrast, a pangenome models the entire set of genomic elements of a given population \citep{Tettelin_2008,cpang2018,Eizenga_2020, Sherman_2020}.
	Pangenomes can be represented as a sequence graph incorporating sequences as nodes and their relationships as edges \citep{Hein1989}.
	%These nodes are shared for identical sequences, such as homologs, paralogs, and orthologs.
	In the variation graph model \citep{Garrison:2018}, genomes are encoded as paths traversing the nodes in the graph.
	%A bidirected graph can even contain both strands of DNA as paths. \\
